---
title: "Homework 2"
author: "Xiangyi Liu (xl3048)"
date: "9/28/2020"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Define the path to the dataset.

```{r}
path_to_data = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```

Read in the dataset Mr.Trashwheel.

```{r read in data}
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```

Read precipitation data! For 2018 and 2017. 
```{r}
precip_2018 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
precip_2017 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Now combine annual precipitation dataframes. In the following code chunk, I create a "helper" tibble that contains pairs of numeric and character ways of representing month, and then merge that (using month number as a key) with the precipitation dataset. This technique is one I use often when I need to recode a moderate or large number of values for a variable. 

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precip_df = 
	bind_rows(precip_2018, precip_2017)
precip_df =
	left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:
* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 2
Defind the path to the dataset.
```{r Define path}
path_to_data_2 = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv"
```

Read in the dataset.
```{r read in}
subway_df =
  read.csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    sep = ",",
    na.strings=c("")
    )%>% 
  janitor::clean_names() %>%
  subset(select = c(line:route11,entry,vending,entrance_type,ada))%>%
  mutate(
    entry = ifelse(entry=="YES",1,2),
    vending =  ifelse(vending=="YES",1,2)
  )
```

The original dataset on NYC Transit has 32 variables. For the homework, we select `r ncol(subway_df)` variables: `r colnames(subway_df)`. The selected variables contains information about subway stations in NYC, including their location, route served, entrance type (stairs/easement), whethere there is vending in the station, and whether it is ADA compliance. 

The variable names of the resulted dataset is cleaned using janitor (all lower-cased, replace space by dash). Variables of route served contains numerous spaces, which are not recognized as NA by R when the dataset is read. Thus, parameter na.string is purposely added to the read.csv function. Further, character vector "entry" and "vending" are converted to logical vector (1=YES, 2=NO).


